### 2023-09-13 README.qmd
Add website stuff, close #1

---
title: CRUSE to Safe Cycling in Ireland
format: gfm
---

<!-- # Introduction

(Lovelace et al., 2017)

• Title page/main information (title, full names and institutional
addresses for all authors, indication of the corresponding author) •
Abstract • Introduction

# Methods

# Results

# Discussion

# Conclusions

# List of abbreviations

# Declarations

## Availability of data and material

## Funding

## Acknowledgements

## Competing interests

## Authors’ contributions

# References

<div id="refs" class="references csl-bib-body hanging-indent"
line-spacing="2">

<div id="ref-lovelace2017" class="csl-entry">

Lovelace, R., Goodman, A., Aldred, R., Berkoff, N., Abbas, A., &
Woodcock, J. (2017). The Propensity to Cycle Tool: An open source online
system for sustainable transport planning. *Journal of Transport and
Land Use*, *10*(1). <https://doi.org/10.5198/jtlu.2016.862>

</div>

</div> -->