### 2023-09-14 README.qmd
Update readme

---
title: CRUSE to Safe Cycling in Ireland
format: gfm
---

The main purpose of this site is to provide broadly-defined documentation in support of the web application, research project, and planning support system hosted at [cruse.bike](https://cruse.bike/).

```{r}
#| echo: false
#| eval: false
# Get the first page with pdftools
remotes::install_cran("qpdf")
qpdf::pdf_subset("_site/paper.pdf", pages = 1:2)
```

See the [paper.qmd](paper.qmd) file for the source code of the paper.