### 2024-03-25 paper.tex
edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{European Transport Research Review}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Cycle Route Uptake and Scenario Estimation (CRUSE): an approach for developing strategic cycle network planning tools},
  pdfauthor={Robin Lovelace; Joey Talbot; Eugeni Vidal-Tortosa; Hussein Mahfouz; Elaine Brick; Peter Wright; Gary O'Toole; Dan Brennan; Suzanne Meade},
  pdfkeywords={Cycling, Open-source, Road Safety, Active Travel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Cycle Route Uptake and Scenario Estimation (CRUSE): an approach
for developing strategic cycle network planning tools}
\author[1]{Robin Lovelace%
\corref{cor1}%
}
 \ead{r.lovelace@leeds.ac.uk} 
\author[1]{Joey Talbot%
%
}

\author[1]{Eugeni Vidal-Tortosa%
%
}

\author[1]{Hussein Mahfouz%
%
}

\author[2]{Elaine Brick%
%
}

\author[3]{Peter Wright%
%
}

\author[2]{Gary O'Toole%
%
}

\author[4]{Dan Brennan%
%
}

\author[4]{Suzanne Meade%
%
}

\affiliation[1]{organization={University of
Leeds},addressline={Institute for Transport Studies, Leeds, LS2 9JT,
United Kingdom},postcodesep={}}
\affiliation[2]{organization={AECOM},addressline={Unit 6, Galway
Technology Park, Parkmore, Galway, Ireland},postcodesep={}}
\affiliation[3]{organization={AECOM},addressline={Winslade House,
Winslade Park, Manor Drive, Clyst St Mary, Exeter, EX5 1FY,
UK},postcodesep={}}
\affiliation[4]{organization={Transport Infrastructure
Ireland},addressline={Parkgate Business Centre, Parkgate Street, Dublin
8, D08 DK10, Ireland},postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper describes an approach for developing strategic cycle network
planning tools. Based on our experience developing and deploying the
Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland, we
outline the underlying methods, including disaggregation of
origin-destination data, incorporation of additional trip purposes,
routing, scenario generation, and development of an intuitive user
interface. Commissioned and used by Transport Infrastructure Ireland,
CRUSE provides estimates of current and potential future cycling levels
under `snapshot' scenarios to inform investment decisions. The publicly
available results at https://cruse.bike/ enable planners, engineers, and
other stakeholders to make more evidence-based decisions. CRUSE goes
beyond previous work by: modeling networks at high spatial resolution;
simulating multiple trip purposes (social, shopping, personal utility,
recreational, and cycle touring) in addition to using official
origin-destination data; and providing estimates of `quietness' (a proxy
for cyclist comfort and route preference) at the route segment level.
Three network types --- `Fastest', `Balanced', and `Quietest' --- help
plan both arterial and residential cycle networks. Workshops and
unstructured interviews with stakeholders show that the tool is already
being used in practice to inform urban, inter-urban, and rural network
planning. The approach is flexible and open-source, allowing the
underlying ideas and code to be adapted, supporting effective cycling
policies and interventions internationally.
\end{abstract}

\begin{keyword}
    Cycling \sep Open-source \sep Road Safety \sep 
    Active Travel
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, sharp corners, enhanced, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden]}{\end{tcolorbox}}\fi

\setstretch{2}
\newpage{}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Transport systems dominated by heavy and powerful private cars are
inefficient, dangerous, and unhealthy. Cars are a major and growing
source of greenhouse gas emissions causing climate change
\citep{winkler2023}, a leading cause of premature death and injury due
to road traffic collisions \citep{globals2018} , and a cause of disease
due to physical inactivity and air, noise and (from tyres) microplastic
pollution \citep{mattsson2023} \citep{welch2023} \citep{cavallaro2024}.
Transport is responsible for 23\% of global greenhouse gas (GHG)
emissions, 70\% of which are from road transport, with passenger cars
accounting for nearly half of transport emissions (around 10\% of global
emissions) \citep{jaramillo2022}. The transport system encourages,
enables and in some cases enforces unsustainable lifestyles. Services
that are only accessible by car lock-in car dependency
\citep{gray2001, shergold2012, motte-baumvol2010}.

Growing evidence of the negative impacts of car-dependent transport
systems has led governments in many countries to set targets and take
actions. In the context of climate, road safety and physicial inactivity
crises, policies to improve transport systems can be classified
according to the `Avoid-Shift-Improve' (ASI) framework
\citep{jaramillo2022}. The ASI framework highlights the importance of
demand reduction (\emph{avoid}ing unnecessary trips), in addition to
mode \emph{shift} to sustainable modes and \emph{improvement} of
existing energy converters, in that order.

Building cycle networks represents a relatively `quick win' within the
context of decarbonization \citep{brand2020} and sustainable mobility
\citep{burns2013}. Although cycling uptake appears on the surface to
only relate to the `shift' part of the ASI framework, closer
consideration of the knock-on impacts of cycling uptake shows that it
can also help avoid unnecessary trips \citep{nello-deakin2020}.
Furthermore, highly efficient ebikes --- which are seeing rapid uptake
in Ireland and other countries --- outperform electric cars, which are
too heavy and expensive. Over-reliance on electric cars could slow the
transition away from car dependency and inadvertently enable ``high
travel lock-in'' \citep{anable2019}. At the European level, the European
Union has a target of reducing GHG emissions by 55\% by 2030, compared
to 1990 levels, and to achieve `net-zero' by 2050 \citep{rosenow2022}.
Climate change mitigation is a major motivation for cycle network plans
\citep{scappini2022}.

Another motivation for cycle network planning at the European level is
the Road Infrastructure Safety Management (RISM) directive (2008/96/EC),
which requires member states to implement a road safety management
system (RSMS) for all public roads. Specifically, ``Member States shall
ensure that the ranking of high accident concentration sections and the
network safety ranking are carried out'' \citep{directiv2008}. Given
that `safety' in this context is usefully quantified as the number of
people killed or seriously injured (KSI) per distance traveled, the
directive requires estimation of distance traveled by mode, down to the
road link level. For active modes, about which there is a paucity of
data compared with motorized modes, this is a major challenge. Better
data to inform road safety policies and interventions is a motivation
for better estimates of `baseline' levels of physical activity at high
levels of geographic resolution \citep{tait2023}.

National governments are increasingly acting on the evidence. In
Ireland, the Road Safety Authority (RSA) has set the target of halving
the number of road traffic deaths and serious injuries by 2030
\citep{national2021}. Doing so while simultaneously enabling rapid
uptake of active modes will require key travel corridors to be
identified `cycle proofed'. Cycling in Ireland represents only 3\% of
total modal share as of the 2016 Census, but accounts for 20\% of
serious injuries and 7\% of all fatalities, driving the development of
cycling policy and infrastructure plans in Ireland. The main frameworks
underpinning these efforts are the Climate Action Plan
\citep{climate2022a}, the National Development Plan, and the National
Roads 2040 strategy \citep{national2023}. These policy frameworks are
backed-up by public support for active travel interventions: while only
3\% of trips in Ireland are made by cycling according to the latest
Census data from 2016 ``83\% of respondents supported enhanced walking
and cycling facilities'' \citep{sustaina2019}.

At the regional level within Ireland, the recently published Greater
Dublin Area Transport Strategy \citep{greater} reinforces these findings
of high support for and capability for cycling: ``nearly a quarter of
adults cycle at least once a week in the Dublin Metropolitan Area'' with
cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around
40\% of the population living in Dublin, suggests that around 150,000
cars could be removed nationwide just by achieving Dublin levels of
cycling in all counties (notwithstanding existing cycling trips and
differences in trip distances). 7 in 10 trips in Ireland are by car
\citep{national} . Cycling has the potential to replace a large
proportion of these trips: ``A high priority must also be given to
cyclists, because trips by this mode have the potential to replace trips
by private car, most specifically for short to medium distance trips,
but increasingly for longer trips as ebikes extend the range of this
mode'' \citep{greater}.

Further evidence of the importance of cycling in Ireland is provided by
the National Strategic Objective (NSO) from the National Development
Plan, which allocates €8.6 billion to sustainable transport
infrastructure including public transport and active travel
interventions. Cycle infrastructure will be developed in synchrony with
the BusConnects project, an entire redesign of the bus network in Dublin
and Cork. It was in this context that Transport Infrastructure Ireland
(TII) commissioned the research reported in this paper, which led to the
Cycle Route Uptake and Scenario Estimation (CRUSE) Tool for Ireland.
Building on previous work, including the Propensity to Cycle Tool (PCT)
for England and Wales, the CRUSE Tool was developed to provide evidence
on current cycling levels and future cycling potential nationwide across
Ireland.

Stakeholder consultation by TII emphasized the need for a strong,
national, systematic but locally-specific evidence base on cycle
networks in both urban and rural Ireland. The evidence base had to scale
nationally to support strategic alignment with national, regional and
local policies, but also be useful for local network planning.

\begin{figure}

{\centering

\includegraphics{images/paste-10.png}

}

\caption{\label{fig-landing}Illustration of the landing page of the
CRUSE Tool for Ireland.}

\end{figure}

This paper describes an approach to strategic cycle network planning
tool development that is open, scalable, and evidence-based. We label
the approach Cycle Route Uptake and Scenario Estimation (CRUSE) and
present a case study of its application in Ireland, the results of which
are publicly available at \href{https://cruse.bike/}{cruse.bike} in the
CRUSE Tool for Ireland, the landing page of which is illustrated in
Figure~\ref{fig-landing}. In Section~\ref{sec-methods}, we outline the
methods used to generate the evidence presented in the CRUSE Tool. In
Section~\ref{sec-results}, we present the results of the CRUSE Tool,
including estimates of current cycling levels and future cycling
potential at the national, regional and local levels. In
Section~\ref{sec-discussion}, we discuss limitations and possible future
improvements to the approach, and the implications of the results for
Ireland before making concluding remarks in
Section~\ref{sec-conclusions}.

\hypertarget{tools-for-estimating-cycling-potential}{%
\section{Tools for estimating cycling
potential}\label{tools-for-estimating-cycling-potential}}

Tools for estimating geographical distribution of cycling potential have
advanced substantially in recent years. Developed by researchers and
developers in academic, public and private sectors, they have evolved
from simple area-based static models tailored to specific regions to
more complex and (in some cases) more generalizable tools. Outputs
include maps with results available at several levels of analysis with
levels of availability ranging from being only available to researchers
and in static maps to open access mapping systems available to the
public. Some of the most relevant existing tools are summarised below
and presented in Table~\ref{tbl-tools}.

\begin{table}

\caption{\label{tbl-tools}\textbf{?(caption)}}\begin{minipage}[t]{\linewidth}
\subcaption{\label{tbl-tools-1}Summary of existing tools for estimating cycling potential.}

{\centering

\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{13em}>{\raggedright\arraybackslash}p{15em}>{\raggedright\arraybackslash}p{12em}>{\raggedright\arraybackslash}p{5em}}
\toprule
Approach & Coverage and input data & Outputs and availability & License\\
\midrule
\cellcolor{gray!6}{Bicycle share model, Parkin et al. (2007)} & \cellcolor{gray!6}{England and Wales, Journeys to work OD census data, at the small-area (wards) level} & \cellcolor{gray!6}{Static tables in academic paper} & \cellcolor{gray!6}{Proprietary}\\
Analysis of  Cycling Potential (ACP), Transport for London (2010, 2016) & London, London Travel Demand Survey & Static tables, graphs, and heatmaps, not publicly available & Proprietary\\
\cellcolor{gray!6}{Prioritization index, Larsen et al. (2013)} & \cellcolor{gray!6}{Montreal, Survey, Road safety, and OD data} & \cellcolor{gray!6}{Map-based heatmap, not publicly available} & \cellcolor{gray!6}{Proprietary}\\
Usage intensity index, Zhang et al. (2014) & Belo Horizonte, Survey, census, and OD data & Static tables and maps, not publicly available & Proprietary\\
\cellcolor{gray!6}{The Cycling Potential Tool (CPT), Phillips and Range (2017)} & \cellcolor{gray!6}{Scotland, Environmental and socioeconomic data, at the small area (output areas) level} & \cellcolor{gray!6}{Maps showing cycling potential in each area} & \cellcolor{gray!6}{Proprietary}\\
\addlinespace
Propensity to Cycle Tool (PCT), Lovelace et al. (2017) & England and Wales, Journeys to work and school OD census data, at the small area (LSOA) level & Online maps, graphs, tables, publicly available at www.pct.bike & Open source\\
\cellcolor{gray!6}{The    Gross Potential for Cycling tool (CPC), Silva et al. (2021 and 2022)} & \cellcolor{gray!6}{21 Portuguese cities, Land use and socio-demographic data, at the small area (census tract) level} & \cellcolor{gray!6}{Static maps showing cycling potential in different areas} & \cellcolor{gray!6}{NA}\\
\bottomrule
\end{tabular}

}

\end{minipage}%

\end{table}

Area-based tools were developed first. Parkin and colleagues presented a
regression model to estimate the percentage of cycling trips to work in
England and Wales based on socioeconomic, transportation, and physical
factors at the small area (wards) level in 2007
\citep{parkin_estimation_2007}. They found that areas with a higher
percentage of females, non-whites, car ownership, lower socioeconomic
classes, income, distance to work, population density, poor highway
conditions, hills, rainfall, and fewer off-road bicycle routes tended to
have a lower proportion of cycling to work.

Three years later, Transport for London created the Analysis of Cycling
Potential (ACP) tool based on a mode choice model that assesses the
likelihood of trips being cycled, with data from the TfL London Travel
Demand Survey as its primary input. The tool generates tables, graphs,
and heatmaps showing the potential for cycling for different trips
purposes in London to guide local cycling initiatives, such as where new
hire `docking stations' should go. The tool does not use
origin-destination data and cannot estimate cycling potential on
specific routes \citep{transport_for_london_analysis_2010}.

Other tools that generate areal data to prioritize investment include a
study that generated a `prioritization index' for Montreal, Canada
\citep{larsen_build_2013}, a `usage intensity index' based on stated
preferences in Belo Horizonte, Brazil, and a Cycling Potential Tool
(CPT) for Scotland \citep{phillips_development_2017}. The CPT is
composed of the base environmental module --- based on eight weighted
factors (population density, hilliness, physical barriers, access to
services, existing cycling mode share, distance to work and school, and
road speed) --- and the quality of service module, which evaluates each
area of interest based on eight factors (surface condition, adjacent
cyclists, comfort factor, conflict, distance between junctions, slope,
access to services, and origin/destination).

A more recent areal-based approach is the designed the Gross Potential
for Cycling (GPC) tool to prioritize areas for cycling infrastructure
and other cycling measures \citep{silva_gross_2021}. The GPC uses two
sets of indicators: population-based indicators (age, potential demand
density, employment density and motorization rate) and area-based
indicators (accessibility, education, public transport, connectivity,
land use mix and relative performance). These indicators are calculated
considering topography, road hierarchy, and average congestion and
presented on a scale from 5 to 10. They are then combined into an
overall score, weighted by their impact on cycling according to the
literature. The ranking of areas by the GPC are
\href{https://boost.up.pt/en/ferramentas/gpc}{publicly available} and
its practical value has been assessed through workshops
\citep{silva_revealing_2022}.

Route-based tools emerged with increasing availability of
origin-destination data, routing engines that can assign trips to
networks, and improvements in computer hardware and software needed to
generate route networks. A prominent example is the the Propensity to
Cycle Tool (PCT), first developed to estimate current and future levels
of cycling at desire line, zone, route, and route network levels for
case study cities \citep{lovelace_mapping_2016}. The approach was
scaled-up to estimate estimate the potential benefits of uptake at zone
and desire line levels nationally, and launched as a publicly available
web application in 2017 \citet{lovelace2017}. Extensions of the PCT
approach have included estimation of benefits at the individual level
\citep{woodcock2018}, addition of travel to school network
\citep{goodman2019}, and improved modeling of impacts on health,
environmental and distributional outcomes \citep{woodcock2021}.
Initially developed just for England, the PCT was extended to cover all
of Wales (for commuter data only) in 2018.

The PCT approach has been applied in other countries, including Ireland
(the topic of this paper), Scotland, and Portugal. In Portugal the
`biclaR' project, based on methods underlying the PCT, has been
developed and deployed for the Lisbon metro region. The resulting
evidence is available in an interactive web application hosted at
\href{https://biclar.tmlmobilidade.pt}{biclar.tmlmobilidade.pt}
\citep{felix2023}. biclaR includes estimates of impacts, using the World
Health Organisation (WHO) `HEAT for Cycling' tool and an `intermodality'
scenario that combines cycling with currently available public transit
options based on General Transit Feed Specification (GTFS) data.

The approach presented in this paper seeks to overcome three limitations
of previous tools to estimate cycling potential: 1) low resolution of
data, with routes starting and ending in administrative zone centroids,
2) limited coverage of trip purposes beyond travel to work and school,
and 3) a web interface that was not user-friendly or intuitive.

\hypertarget{sec-methods}{%
\section{Methods and data}\label{sec-methods}}

\hypertarget{sec-disaggregation}{%
\subsection{Disaggregation of origin-destination
data}\label{sec-disaggregation}}

A feature of active travel interventions is that they require dense
networks of routes to be effective \citep{parkin2018}. This means that
high levels of geographic resolution are needed in the data used to
estimate cycling potential. However, datasets on travel patterns are
often only available at the level of administrative zones. The Central
Statistics Office (CSO) in Ireland provides Place of Work, School or
College Census of Anonymized Records
(\href{https://www.cso.ie/en/census/census2016reports/powscar/}{POWSCAR})
data on the number of people traveling to work and school at the
Electoral Division (ED) level, for example.

In the PCT, the method used to convert OD data to route networks was to
calculate a single route between the population weighted centroids of
the zones associated with each OD pair. This method works fine when the
OD data represents movement between small areas, but was not appropriate
for generating route networks from the POWSCAR data because zone
centroids are so far apart that the resulting route networks would be
sparse and unrealistic. To tackle this issue we developed a new method
for OD data disaggregation called `jittering' \citep{lovelace2022}. The
method works by first disaggregating the OD data based on a
`disaggregation threshold' and then assigning each disaggregated
`sub-OD' pair to `subpoints' within each zone. As shown in
Figure~\ref{fig-dublin}, the resulting route networks are dense, even in
rural areas.

\begin{figure}

{\centering

\includegraphics{images/paste-8.png}
\includegraphics{images/paste-9.png}

}

\caption{\label{fig-dublin}Illustration of the density of route networks
generated by the CRUSE Tool for Dublin City and surroundings (top) and
rural County Mayo on the West coast of Ireland (botton). Source: CRUSE
Tool, publicly available at \href{https://cruse.bike/}{cruse.bike}.}

\end{figure}

\hypertarget{sec-trip-purposes}{%
\subsection{Additional trip purposes}\label{sec-trip-purposes}}

A limitation of the original PCT was that it only included travel to
work data. This was partially addressed by the inclusion of travel to
school based on data from the Department for Education in England
\citep{goodman2019}. An advantage of the POWSCAR OD data over OD
datasets derived from census surveys in many countries is that it
includes travel to school data. We commissioned a version of POWSCAR
that included a breakdown of the total flows between OD pairs by purpose
and mode, enabling a more realistic estimation of the `Baseline' cycling
network.

However, travel to school and work only constitute around only 30\% of
all trips in Ireland. To overcome this issue, we developed a spatial
interaction modeling methodology to estimate the number of trips between
each OD pair for additional trip purposes. The classification of trip
purposes used in the CRUSE Tool was guided mainly by the trip purpose
classification found within the National Household Travel Survey (NHTS),
but with the addition of categories based on the comprehensive POWSCAR
data, and the need to include recreational trips and multi-stage trips
(not yet implemented). An overview of the trip purposes used in CRUSE is
presented in Figure~\ref{fig-trip-purposes}.

\begin{figure}

{\centering

\includegraphics{images/paste-4.png}

}

\caption{\label{fig-trip-purposes}Trip purposes used in the CRUSE Tool.}

\end{figure}

\hypertarget{routing}{%
\subsection{Routing}\label{routing}}

Routes in CRUSE are generated by CycleStreets, a not-for profit
transport consultancy and web development company who provide
application programming interfaces (APIs) supplying a range of datasets
for cycle planning and advocacy. Although based in Cambridge, UK,
CycleStreets provide routing services `by cyclists, for cyclists'
internationally, including in Ireland. The CycleStreets routing engine
is based on OpenStreetMap (OSM) data, which is continuously updated by a
global community of volunteers.

While CycleStreets offers a free routing API, we commissioned a custom
routing service to enable:

\begin{itemize}
\tightlist
\item
  Calculation of hundreds-of-thousands of routes, which is beyond the
  terms of service of the free API.
\item
  Making changes to the routing profiles, including allowing routing on
  trunk roads, which are sometimes avoided in the default routing
  profiles.
\item
  Control over the version of OSM data being used for the routing,
  allowing regular updates to the route networks as OSM data is updated.
\end{itemize}

We calculated three network types for each disaggregated (`jittered') OD
pair: `Fastest', `Balanced' and `Quietest'. As outlined on the
CycleStreets'
\href{https://www.cyclestreets.net/help/journey/howitworks/}{website},
the `Fastest' routes minimize journey time, accounting for traffic
lights and surface type. The `Quietest' route type minimizes busy
sections of road, allowing high `diversion factors' from the fastest
route on quieter but less direct ways (often avoiding roads and
interactions with motor vehicles altogether where possible). The
`Balanced' route type is a compromize between the two, minimising
journey time while avoiding the busiest roads. Allowing users to switch
between these three route types enables them to consider the trade-offs
between directness and `cycle friendliness' when planning new
infrastructure, as shown in Figure~\ref{fig-route-types}.

These map are available on the `Route types' page for each county (see
\href{https://cruse.bike/kildare/route-types}{cruse.bike/kildare/route-types}).
In addition to showing the estimated routes under each scenario, the
page also presents summary statistics on the network:

\begin{itemize}
\item
  On the fastest network, 25\% of the distance cycled occurs in
  non-hostile segments and 7\% in cycle-friendly segments.
\item
  Under the baseline scenario, 52\% of the distance cycled on the
  quietest network occurs in non-hostile segments and 15\% in
  cycle-friendly segments.
\end{itemize}

These statistics can be revealing: in Kildare it suggests that around
half of all cycling activity occurs on network segments that are hostile
(with a high inferred level of traffic stress), even when efforts are
taken to avoid busy segments. For the Fastest network, which may be more
realistic for utility cycling, the proportion of cycling on hostile
segments is even higher at 75\%.

\begin{figure}

\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_quietest_godutch.png}

}

}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/naas_fastest_godutch.png}

}

}

\end{minipage}%
\newline
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_quietest_godutch.png}

}

}

\end{minipage}%
%
\begin{minipage}[t]{0.50\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics[width=2.60417in,height=\textheight]{images/newbridge_fastest_godutch.png}

}

}

\end{minipage}%

\caption{\label{fig-route-types}Illustration of the Quietest (left) and
Fastest (right) route networks for Naas (top) and Newbridge (bottom).}

\end{figure}

\hypertarget{sec-scenarios}{%
\subsection{Scenarios of cycling potential}\label{sec-scenarios}}

The datasets outlined above were used to generate estimates of cycling
currently (the `Baseline' scenario) and under four scenarios of cycling
potential: Near Market, Climate Action Plan, Go Dutch, and Ebike. Each
of these scenarios is described below, as illustrated in
Figure~\ref{fig-scenarios}.

\begin{figure}

{\centering

\includegraphics{images/scenarios_kildare.png}

}

\caption{\label{fig-scenarios}Distance-mode share graph for Kildare used
to communicate the different scenarios. Source: county level view for
Kildare in the open access CRUSE Tool at
\url{https://cruse.bike/kildare}.}

\end{figure}

\hypertarget{baseline}{%
\subsubsection{Baseline}\label{baseline}}

The Baseline scenario approximates current cycling levels. As outlined
in Section~\ref{sec-trip-purposes}, the Baseline scenario includes
travel to work and school, and additional trip purposes. Cycling levels
were taken from the POWSCAR data, which includes the number of people
traveling by each mode between each OD pair.

\hypertarget{near-market}{%
\subsubsection{Near Market}\label{near-market}}

The Near Market scenario approximates the level of cycling that would be
achieved if levels of cycling uptake observed in areas of Ireland with
high levels of cycling according to the 2016 Census were achieved
everywhere, accounting for differences in trip distances and hilliness
levels. The scenario is implemented as follows:

\begin{itemize}
\tightlist
\item
  Calculate distance decay curves for Dublin for the base year (2016,
  using POWSCAR data) by fitting a model to the relevant OD data after
  it has been converted to a route network dataset
\item
  Apply the Near Market model to the hilliness and distance values for
  each county during the build process
\item
  Add the current level of cycling to the Near Market model
\end{itemize}

\hypertarget{climate-action-plan}{%
\subsubsection{Climate Action Plan}\label{climate-action-plan}}

The Climate Action Plan scenario models the transport emissions
reductions targeted in the Irish Government's
\href{https://www.gov.ie/en/publication/6223e-climate-action-plan-2021/}{Climate
Action Plan 2021}, which aims for a 51\% cut in overall greenhouse gas
emissions by 2030 as part of the pathway to net-zero emissions by 2050.
For transport, this includes 500,000 extra walking, cycling and public
transport trips per day by 2030. In terms of car travel, the target is
to ``Increas{[}e{]} the proportion of kilometers driven by passenger
electric cars to between 40 and 45\% by 2030, in addition to a reduction
of 10\% in kilometers driven by the remaining internal combustion engine
cars.'' This equates to a 5.5 to 6\% reduction in total car km driven.

To model this decrease in car driver km, cycling uptake increases in
line with the Go Dutch scenario. However, we only model shift from
driving to cycling. There is no shift from other modes of transport to
cycling.

\hypertarget{go-dutch}{%
\subsubsection{Go Dutch}\label{go-dutch}}

Under the Go Dutch scenario cycling reaches levels equivalent to those
found in the Netherlands, taking account of the effects of route
hilliness (measured as mean gradient) and route distance. This scenario
uses the same model that was used PCT, allowing trips to shift from any
other mode to cycling \citep{lovelace2017}.

\hypertarget{ebike}{%
\subsubsection{Ebike}\label{ebike}}

Also based on the PCT scenario with the same name, the Ebike scenario in
CRUSE takes Go Dutch cycling uptake, and adds onto this the impact of
increased ebike usage, which allows for longer cycle trips. However,
travel to primary and secondary schools still uses Go Dutch uptake,
since no ebike scenario has been developed for school journeys, and
children may be less likely to own ebikes than adults.

\hypertarget{sec-ui}{%
\subsection{User interface}\label{sec-ui}}

The CRUSE web application is statically hosted, meaning that it can be
accessed without the need for a server running software such as the R
package `shiny' or Python packages such as `streamlit' or `flask' in the
background \citep{wickham2021}. Typical intended user stories are
illustrated in Figure~\ref{fig-user-stories}, which shows that the tool
is designed to be used by both professional and non-professional users.
For the main target audience, professional transport planners working at
the county level, the tool provides a range of outputs, including
estimates of cycling potential at the county and network level. The
provision of balanced (the default), quietest and fastest route networks
enables planners to consider the trade-offs between directness and
`cycle friendliness' when planning new infrastructure. The tool also
provides data downloads, enabling estimates of cycling potential on the
network to be visualised and analysed with other tools such as QGIS,
Python or R.

For non-professional users, the tool provides a simple interface to
explore the cycling potential of the network. By providing a single
landing page that is suitable for both professional and non-professional
users (such as an advocate or parent interested in safe routes to
school), the tool aims to facilitate communication between these groups.

\begin{figure}

{\centering

\includegraphics{images/mermaid-figure-1.png}

}

\caption{\label{fig-user-stories}Hypothetical user stories for the CRUSE
Tool illustrating that it can be used in different ways by different
user groups.}

\end{figure}

\hypertarget{sec-results}{%
\section{Results}\label{sec-results}}

The main result of the work presented in this paper is an open access
web application and datasets on current and potential future cycling
levels in Ireland, with evidence provided at county and route segment
levels. Like the PCT on which CRUSE builds, a key feature of the results
is that they are systematic and available for all counties, and most
routes on which cycling is permitted, in Ireland. The full set of
results are therefore too extensive to present in their entirety in this
paper. Through the interactive web application, users can explore the
data to generate the results that are most relevant to their needs,
whether that is finding the cycling potential on a particular road or
finding `weak links' or barriers in the cycle network associated with a
particular school, work place or other destination.

Instead of trying to show such use cases, of which there are many
hundreds, we present a selection of results to illustrate the main
features of the CRUSE Tool.

To illustrate the results in urban areas, we took the top 4 cities in
Ireland by population: Dublin City, Cork, Limerick and Galway (with city
populations of 0.6, 0.2, 0.1 and 0.1 million respectively). In
Figure~\ref{fig-city-results} we present the cycle route networks within
a 3km radius circle centred on each of these cities.

\begin{figure}

{\centering

\includegraphics{images/rnet_types.png}

}

\caption{\label{fig-city-results}Fastest, balanced and quietest cycle
networks within 3km circular areas containing the four largest cities in
Ireland. The networks are coloured by cycle friendliness, with greener
segments representing more cycle friendly segments. Widths vary in
proportion to the potential level of cycling under the `Go Dutch'
scenario.}

\end{figure}

Aside from the widely varying sizes and route networks of each city,
with Dublin having by far the densest network, it is clear from the
visualisation of cycling friendliness that there are many gaps in the
networks. Dublin has had received most investment in cycling
infrastructure and this is apparent from the high proportion of the
route network that is green, even in the `Fastest' route network
representing routes that prioritize speed and directness over comfort
and safety. Dublin also has the highest mode share of cycling of the
four cities, adding to the evidence base showing that investment in
cycling infrastructure is effective in increasing cycling uptake.

Although Dublin City has made progress, there are still many parts of
the Fastest route network, and even some parts of the Balanced and
Quietest route networks, that are not cycle friendly and which have high
cycling potential. According to a recent report, 71\% of residents in
the Dublin Metropolitan Area support ``more cycle tracks along roads,
physically separated from traffic and pedestrians'' \citep{walking2021}.
The results for the Dublin area could help prioritise investment in
cycling infrastructure in the city.

\hypertarget{sec-discussion}{%
\section{Discussion}\label{sec-discussion}}

The aim of the paper was to describe the design, features and potential
use of the `CRUSE' approach and resulting tool. As outlined in the
previous section, the CRUSE Tool provides evidence on current and
potential future cycling levels across Ireland down to the street level,
with potential assigned to `fastest', `balanced' and `quietest' route
networks. This provision of multiple scenarios of behavior change
\emph{and} multiple scenarios of investment, for example in cycling
infrastructure next to major roads vs quiet residential streets, is a
key feature of the CRUSE Tool.

\hypertarget{uses-of-the-tool-in-practice}{%
\subsection{Uses of the tool in
practice}\label{uses-of-the-tool-in-practice}}

As highlighted in a paper on cycling infrastructure preferences based on
a case study of Dublin, both directness and quietness are important
\citep{caulfield2012}:

\begin{quote}
Direct routes with short journey times were found to be the most
important positive variable for existing cyclists and non-cyclists in
determining route choice. This is followed by infrastructure type, the
number of junctions along the route, traffic speed and cyclist volumes.
In terms if infrastructure, regardless of the level of cycling
confidence, routes which have `no facilities' or `bus/cycle lanes' are
the least favoured cycle route types.
\end{quote}

The CRUSE Tool can help both in terms of describing the current
situation, and also prioritise investment in those direct routes with
high cycling potential that lack adequate cycling infrastructure
\citep{caulfield2012}. It will also support reporting for the RISM
Directive on road casualties as part of the road safety management
system.

\hypertarget{comparison-with-other-tools-and-contribution-to-the-field}{%
\subsection{Comparison with other tools and contribution to the
field}\label{comparison-with-other-tools-and-contribution-to-the-field}}

While the CRUSE Tool is not the first national and publicly available
tool for cycling planning, it has some key features that make it
relevant for other countries, regions and road authorities tasked with
making their transport systems safe and sustainable:

\begin{itemize}
\tightlist
\item
  The results are \emph{open access} meaning that any stakeholder in the
  planning system can access the evidence. This will help to democratize
  the transport planning process and make wider conversations about
  transport planning more evidence-based and less polarized
  \citep{lovelace2020}, something that is particularly relevant given
  the potentially polarizing nature of pro-cycling interventions
  \citep{wild2017}.
\item
  The results are fully reproducible (code to be released pending
  sign-off by TII's IT team), preventing `cloud lock in' to a
  potentially monopolistic consultancy, and encouraging input from the
  wider open source community \citep{lovelace2021, dhir2017}.
\item
  By covering a wide range of trip purposes --- not just travel to work
  \citep{lovelace2017, heinen2010} and travel school \citep{goodman2019}
  as covered in previously published research on open access national
  cycle network planning tools --- the results capture a higher
  proportion of cycling potential including key rural trips which are
  often under-represented in models of active travel.
\end{itemize}

As the benefits of strategic planning for cycling become more apparent
\citep{scappini2022} we expect the demand for tools such as CRUSE to
grow. However, the approach is not without limitations and does not meet
all requirements in Ireland or indeed any country where the methods are
applied.

A limitation of the approach as implemented in the CRUSE tool is that
the results are `static', reducing its ability to support monitoring of
growth in cycling, upgrades of existing cycling infrastructure and
dynamic (year-by-year) exposure information to estimate crash rates.
Exposure information is a key part of the the European RISM Directive,
for reporting collision rates of vulnerable road users by 2024. These
factors mean that the tool can be seen as an open access `leverage
point', providing key information for and supporting ambitious plans in
many aspects the planning system, from network design to post-build
monitoring \citep{lovelace2020}.

\hypertarget{sec-conclusions}{%
\section{Conclusions}\label{sec-conclusions}}

The CRUSE Tool is an open access web application for strategic cycle
network planning and prioritization of road safety interventions across
Ireland. Building on previous work, it provides a nationally consistent
evidence base, providing valuable insights to planners and other
stakeholders in the transport planning process, at national and local
levels. Because the results are available at the route segment level,
the tool can be used to identify `weak links' in the cycle network
\citep{vybornova2022}, and to prioritize investment in cycling
infrastructure to maximize health, equality, and other benefits
\citep{mahfouz, woodcock2021}. Furthermore, the provision of the
evidence in a free and publicly available website, hosted at
\href{https://cruse.bike}{cruse.bike}, means that it can be used by
anyone, encouraging wider participation and more evidence-based debate
about transport planning.

A key feature of the project methodologically is its calculation of
current and future potential not only for travel to work and travel to
school, but also for other trip purposes, including recreational trips.
This required the development of spatial interaction models and
estimation of the relative attractiveness of different destinations for
different trip purposes, an area of active research where new
developments could be incorporated into the tool in future
\citep{hasova2022}.

The CRUSE Tool and the underlying research and methods are not without
limitations, suggesting future areas of research, data collection needs,
and policy application. The route network level results have not been
validated to the extent we had planned at the outset of the project. We
compared route network level results with `ground truth' data from cycle
counter datasets across Dublin to test different network types and to
ensure that additional trip types added to work and educational trips
increased the quality of fit under the baseline scenario. However, the
size of the counter network was insufficient to provide an opportunity
for robust evaluation of model performance, suggesting a combined
program of new count data collection and analysis should be a priority
for future work, with findings feeding directly into work to improve the
outputs of the the CRUSE Tool.

More broadly, the tool's outputs are limited to just one mode of active
travel (cycling), ignoring walking and wheeling, including wheelchair
use and a range of wheeled devices such as scooters, bike trailers and
e-cargo bikes that can be used to escort children to school. Sustainable
transport policies should plan for walking, wheeling and cycling, and
there are many co-benefits of broadly defined active travel
interventions that benefit all active modes, such as measures to reduce
heavy motor traffic speeds and volumes in areas and along corridors with
high active travel potential. This raises the question of whether other
active modes should be incorporated into the results using the OD-based
approach outlined in this paper, or whether different modeling
approaches are needed to properly capture the shorter distance trips
typically made by walking \citep{cooper2018}. Furthermore, the estimates
of cycling presented in the CRUSE Tool omit multi-modal trips including
public transport, and omit trip chaining, due to the need to capture a
large portion of cycling potential within the resource constraints of
the project.

The CRUSE Tool is already used in practice to support more ambitious and
data-driven planning for safe cycling routes in multiple counties across
Ireland. We hope that the underlying approach, and the publicly
available evidence provided in the CRUSE Tool itself, provide a basis
for future reproducible research, open source software development into
strategic cycle network planning tools in Ireland and other countries.
Active travel represents a win-win-win for physical activity,
environment, and local economic opportunities. In combination with
broader sustainable mobility measures and policies to reduce motor
traffic speeds and volumes, tools such as CRUSE can support the fast and
fair decarbonisation of transport systems worldwide.

\hypertarget{list-of-abbreviations}{%
\section{List of abbreviations}\label{list-of-abbreviations}}

APIs: Application Programming Interfaces

CSO: Central Statistics Office

CRUSE: Cycle Route Uptake and Scenario Estimation

ED: Electoral Division

GHG: Greehouse Gas

GTFS: General Transit Feed Specification

KSI: Killed and Seriously Injured

NRA: National Road Authority

NRN: National Road Network

OD: Origin-Destination, typically referring to origin-destination data
which contains information on the number of people traveling between
each pair of zones

OSM: Open Street Map

PCT: Propensity to Cycle Tool

POWSCAR: Place of Work, School or College Census of Anonymized Records

TII: Transport Infrastructure Ireland

\hypertarget{declarations}{%
\section{Declarations}\label{declarations}}

\hypertarget{availability-of-data-and-material}{%
\subsection*{Availability of data and
material}\label{availability-of-data-and-material}}
\addcontentsline{toc}{subsection}{Availability of data and material}

Data was obtained from the Central Statistics Office (CSO) and Transport
Infrastructure Ireland (TII) under license and cannot be shared
publicly. The code used to generate the results presented in this paper
is fully reproducible and available at {[}to be added post review.{]}

\hypertarget{funding}{%
\subsection*{Funding}\label{funding}}
\addcontentsline{toc}{subsection}{Funding}

This work was funded by Transport Infrastructure Ireland (TII).

\hypertarget{acknowledgements}{%
\subsection*{Acknowledgements}\label{acknowledgements}}
\addcontentsline{toc}{subsection}{Acknowledgements}

{[}To be added post review.{]}

  \bibliography{bibliography.bib,cruse.bib}

\end{document}